Analyzing the Effect of Surface in Tennis Grand Slams

### 2018-09-14 contest/surface-type/paper-tennis.Rmd
spell check lol

We use models that take into account both individual and group effects such as in the Gaussian-process player production basketball model or predicting individual soccer performance  [@page2013; @egidi2018].  Both of those models had success using hierarhcical Bayesian models, which we employ in our own models. More specifically, we model the players' expected points in a match based on the player's own characteristics, the court/tournament effects, and the oponent's ranking.

@@

We first examine the distibution of points earned per match. We find that the distribution is similar across tournament, with Wimbledon differing slightly from the other grand slams. As expected, there are more points earned in the WTA than the ATP due to the differing numbers of games played. Also unsuprisingly, the winners of the match tended to earn more points than the losers.

@@

It is also commonly thought that Spaniards play better on clay. We are intersted in whether Spaniards win the French Open more than they win in other tournaments. It does appear that Spaniards are winning the French Open more than they are winning other tournaments. But, this result is not significant. In addition, the ranking of Spaniards in the French Open is, on average, higher than other tournmanents, which may help explain this common misconseption, although this result is not significant either.

>>>>

We use models that take into account both individual and group effects such as in the Gaussian-process player production basketball model or predicting individual soccer performance  [@page2013; @egidi2018].  Both of those models had success using hierarchical Bayesian models, which we employ in our own models. More specifically, we model the players' expected points in a match based on the player's own characteristics, the court/tournament effects, and the opponent's ranking.

@@

We first examine the distribution of points earned per match. We find that the distribution is similar across tournament, with Wimbledon differing slightly from the other grand slams. As expected, there are more points earned in the WTA than the ATP due to the differing numbers of games played. Also unsurprisingly, the winners of the match tended to earn more points than the losers.

@@

It is also commonly thought that Spaniards play better on clay. We are interested in whether Spaniards win the French Open more than they win in other tournaments. It does appear that Spaniards are winning the French Open more than they are winning other tournaments. But, this result is not significant. In addition, the ranking of Spaniards in the French Open is, on average, higher than other tournaments, which may help explain this common misconception, although this result is not significant either.
